---
title: "Function_NHANES"
author: "S.H.Hosseini"
date: "30/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Function to download the NHANES data

This function can be used to download all NHANES datasets for the years of your choice. download_nh() function has 3 arguments that are data_name, year and name_to_label. 

by default the function download the selected datasets since 1999-2000. 

by default the variables names are not converted to variables' labels -> so one would like to use the variables labels instead of their names, the argument name_to_label should change be set as TRUE.

data_name allows us to choose the datasets of our choice. For instance, if one requires the datasets containing information about insulin for years 2001-2002 and 2003-2004 the function can be used as follows: 

download_nh(data_name = "insulin",year = c("2001-2002","2003-2004"))

one can download more than one dataset if it is required. The data_name argument accept regex so you have to use "|" to distinguish between the datasets inside data_name argument to get more than one dataset. For instance, if you require datasets including information about body measures and blood pressure you could use the function as follows: 

download_nh(data_name = c("blood pressure|body measure"),year = c("1999-2000","2017-2018"))


# FUNCTION to RUN 
```{r}

download_nh <- function(data_name=NULL,year=c("1999-2000","2001-2002","2003-2004","2005-2006", "2007-2008", "2009-2010", "2011-2012", "2013-2014","2015-2016","2017-2018"),name_to_label=FALSE) {

library(haven)
library(tidyverse)
library(janitor)
library(glue)
library(rvest)
library(stringr)
library(data.table)
  
file_list3 <- read_csv("https://raw.githubusercontent.com/economer/NHANES/master/file_list3.csv",progress = F) %>%
  mutate(file_name = tolower(file_name))


name_list3 <- read_csv("https://raw.githubusercontent.com/economer/NHANES/master/name_list3.csv",progress = F)

## function to download xpt files 
 import_xpt <- function(links) {
  df <- haven::read_xpt(links)
 }

# use str_detect to filter the files required
selected_data <-  file_list3 %>%
  filter(str_detect(string = file_name,pattern = data_name),
         Year == {{year}}
         ) 
# download the data selected 
selected_data_map <- selected_data %>%
  mutate(data_files = map(download_links,import_xpt))

## select year and data files
selected_data_map1 <- selected_data_map %>%
  select(Year,file_name, data_files) 

## convert the nested data frames to a list
## 
selected_data_map_list <- selected_data_map1 %>%
  map(~list(.))

## keep the datasets' lists 
selected_data_map_list <- selected_data_map_list[["data_files"]][[1]]

## Join the datasets by SEQN

final_dataset <- plyr::join_all(dfs = selected_data_map_list,by = "SEQN",type = "full")

if(name_to_label == TRUE) {
  
  name_list3 <- name_list3 %>%
  distinct(var_name,.keep_all = T)
  setnames(x = final_dataset,old = as.character(name_list3$var_name),new =
           as.character(name_list3$var_label),skip_absent = T)
  final_dataset <- final_dataset %>%
    as_tibble(.name_repair = "minimal")

final_dataset

} else {
    final_dataset
}
  
} 




```

## Examples

```{r}
## NOTE: if you need to know the name of the datasets available for each year, you could run the following code and then use the names inside the data_name argument as instructed above. 
file_list3 <- read_csv("https://raw.githubusercontent.com/economer/NHANES/master/file_list3.csv",progress = F) %>%
  mutate(file_name = tolower(file_name))

file_list3 %>%
  group_by(Year, file_name) %>%
  count()

## 
# download insulin datasets for 1999-2002
# 
insulin_1999 <- download_nh(data_name = "insulin" ,year = "1999-2000",name_to_label = TRUE)

## download blood pressure and body measure datasets for 1999-2000 and 2017-2018
 
blood_pressure_body_measure <- download_nh(data_name = c("blood pressure|body measure"),year = c("1999-2000","2017-2018"))
```

